---
title: "Spotify Hit Predictor"
author: "Gino Assenmacher, Nathan Koenig, Johanna Langemeyer, Lukas Resch, Ben Roemer"
date: "2022-10-27"
output:
  html_document: 
    highlight: tango
    theme: united
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About the dataset
The dataset is called 'The Spotify Hit Predictor Dataset (1960-2019)' and consistes of features for tracks fetched using Spotify's web API. For the sake of the analysis we will focus on the hits from 2010-2019 as we believe that the factors that predict a hit song changed over time. Therefore, more recent data allows for better predictions.
The variable that we are trying to predict is weather the song was a hit ('1') or a flop ('0'). This classification is determined by the following criteria for a flop:

* The track must not appear in the 'hit' list of that decade.
* The track's artist must not appear in the 'hit' list of that decade.
* The track must belong to a genre that could be considered non-mainstream and / or avant-garde. 
* The track's genre must not have a song in the 'hit' list.
* The genre list for the particular decades are as follows:
* The track must have 'US' as one of its markets.

The dataset contains the following data:
```{r echo = FALSE}
spot10 <- read.csv("dataset-of-10s.csv")
str(spot10)
```
The dataset can be found [here.](https://www.kaggle.com/datasets/theoverman/the-spotify-hit-predictor-dataset)

## Business Case
Large record labels are flooded by new artists and requests for releasing new songs. With this data we can build an algorithm that predicts weather a song will become a hit or not. By doing so, we can facilitate investment decision in e.g. marketing for certain song. To quantify this, we will use an estimate of the costs of marketing a song and on the other hand the returns a hit song can bring.

## Cleaning the Dataset
There is not a lot of cleaning needed as the dataset is complete and pre-cleaned. We will just leave out certain variables from the regression.

```{r}

spot10 <- read.csv("dataset-of-10s.csv")
spot00 <- read.csv("dataset-of-00s.csv")
spot90 <- read.csv("dataset-of-90s.csv")
spot80 <- read.csv("dataset-of-80s.csv")
spot70 <- read.csv("dataset-of-70s.csv")
spot60 <- read.csv("dataset-of-60s.csv")

spot10$uri <- NULL
spot00$uri <- NULL
spot90$uri <- NULL
spot80$uri <- NULL
spot70$uri <- NULL
spot60$uri <- NULL

spot10$key <- as.factor(spot10$key)
spot00$key <- as.factor(spot00$key)
spot90$key <- as.factor(spot90$key)
spot80$key <- as.factor(spot80$key)
spot70$key <- as.factor(spot70$key)
spot60$key <- as.factor(spot60$key)

spot10$mode <- as.factor(spot10$mode)
spot00$mode <- as.factor(spot00$mode)
spot90$mode <- as.factor(spot90$mode)
spot80$mode <- as.factor(spot80$mode)
spot70$mode <- as.factor(spot70$mode)
spot60$mode <- as.factor(spot60$mode)

spot10$decade <- as.factor("2010")
spot00$decade <- as.factor("2000")
spot90$decade <- as.factor("1990")
spot80$decade <- as.factor("1980")
spot70$decade <- as.factor("1970")
spot60$decade <- as.factor("1960")

spotify <- rbind(spot10, spot00, spot90, spot80, spot70, spot60)

spotify$track <- NULL
spotify$artist <- NULL

```

```{r}

set.seed(123)
spotify_random <- spotify[sample(nrow(spotify)),]

spotify_train <- spotify_random[1:(41106/2),] 

spotify_test <- spotify_random[((41106/2)+1):nrow(spotify),]


normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

spotify_train_dummy <- as.data.frame(model.matrix(~.-1,spotify_train))
spotify_test_dummy <- as.data.frame(model.matrix(~.-1,spotify_test))

spotify_train_norm <- as.data.frame(lapply(spotify_train_dummy, normalize))
spotify_test_norm <- as.data.frame(lapply(spotify_test_dummy, normalize))

write.csv(spotify, "C:/Users/gasse/Documents/School/to414/gp2/spotify.csv", row.names = FALSE)
write.csv(spotify_random, "C:/Users/gasse/Documents/School/to414/gp2/spotify_random.csv", row.names = FALSE)
write.csv(spotify_train, "C:/Users/gasse/Documents/School/to414/gp2/spotify_train.csv", row.names = FALSE)
write.csv(spotify_test, "C:/Users/gasse/Documents/School/to414/gp2/spotify_test.csv", row.names = FALSE)
write.csv(spotify_train_dummy, "C:/Users/gasse/Documents/School/to414/gp2/spotify_train_dummy.csv", row.names = FALSE)
write.csv(spotify_test_dummy, "C:/Users/gasse/Documents/School/to414/gp2/spotify_test_dummy.csv", row.names = FALSE)
write.csv(spotify_train_norm, "C:/Users/gasse/Documents/School/to414/gp2/spotify_train_norm.csv", row.names = FALSE)
write.csv(spotify_test_norm, "C:/Users/gasse/Documents/School/to414/gp2/spotify_test_norm.csv", row.names = FALSE)

```